---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = TRUE
)
```

# CalibratedVA

<!-- badges: start -->
  [![Travis build status](https://travis-ci.com/jfiksel/CalibratedVA.svg?branch=master)](https://travis-ci.com/jfiksel/CalibratedVA)
  <!-- badges: end -->


## Installation

Before installation of CalibratedVA, we recommend first installing the latest version
of LaplacesDemon from [GitHub](https://github.com/):

```r
# install.packages("remotes")
remotes::install_github("LaplacesDemonR/LaplacesDemon")
```

You can then install the development version of CalibratedVA from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("jfiksel/CalibratedVA")
```
## Getting started with CalibratedVA

### PHMRC Data

The Population Health Metrics Research Consortium (PHMRC) study contains GS COD data
for children, neonates and adults in 4 countries, making it a perfect dataset
to demonstrate the merits of `CalibratedVA`. 

We have used the openVA package to obtain cause-of-death predictions for adult
deaths from Tanzania. Both InSilicoVA and Tariff were trained using deaths
from the other 3 countries. We obtained individual cause probability predictions
from InSilicoVA and individual top cause-of-death predictions from Tariff. We
will also read in the GS-COD for each death from Tanzania.

```{r load_phmrc}
library(CalibratedVA)
data("insilico_tanzania")
data("tariff_tanzania")
data("gs_cod_tanzania")
```

There are `r ncol(insilico_tanzania)` causes, which is too many causes for 
CalibratedVA to be useful. We will map these causes to broader causes, using the
following cause map.

```{r load_cause_map}
data("phmrc_adult_cause_map")
```

To map the causes, we can provide either the probability matrix:

```{r map_causes_insilicova}
insilico_tanzania <- map_causes(insilico_tanzania, phmrc_adult_cause_map)
```

Or the character vector with the predicted cause or GS-COD for each individual

```{r map_causes_tariff}
tariff_tanzania <- map_causes(tariff_tanzania$cause, phmrc_adult_cause_map)
gs_cod_tanzania <- map_causes(gs_cod_tanzania, phmrc_adult_cause_map)
```


### Single algorithm predictions

We will first run CalibratedVA using the Tariff predictions. To run CalibratedVA, we will first select a set of individuals to be in our "hospital" set--that is, individuals for whom we know their GS-COD. These
individuals will be used for estimation of algorithm misclassification rates. We will
simply use the first 200 individuals from the data. We will use the default settings
for all hyperparameter values.


```{r tariff_calibratedva}
causes <- colnames(tariff_tanzania)
tariff_calibratedva <- calibratedva(A_U = tariff_tanzania[-(1:200),],
                                    A_L = tariff_tanzania[1:200,],
                                    G_L = gs_cod_tanzania[1:200,],
                                    causes = causes)
```

We can use the `ggmcmc` package to obtain CSMF estimates for the individual causes,
along with credible intervals.

```{r tariff_csmf}
library(ggmcmc)
P <- data.frame(
  Parameter=paste0("p[", 1:length(causes), "]"),
  Label=causes)
tariff_csmf_samples <- ggs(tariff_calibratedva$samples, par_labels=P, family="p")
### CSMF with credible intervals
tariff_csmf <-
  tariff_csmf_samples %>%
  group_by(Parameter) %>%
  summarise(csmf = mean(value),
            ci_L = quantile(value, .025),
            ci_U = quantile(value, .975))
tariff_csmf 
```

We can also view the traceplot

```{r tariff_csmf_traceplot, fig.width = 8, fig.height = 12}
ggs_traceplot(tariff_csmf_samples)
```

And a posterior density plot

```{r tariff_csmf_density_plot, fig.width = 8, fig.height = 12}
ggs_density(tariff_csmf_samples)
```

### Using predictions from multiple algorithms (ensemble)

We can use both the predictions from both InSilicoVA and Tariff to 
better estimate the CSMF using the ensemble approach. We simply need to format
the predictions into a list of matrices

```{r format_for_ensemble}
ensemble_unlabeled <- list(tariff_tanzania[-(1:200),], insilico_tanzania[-(1:200),])
ensemble_labeled <- list(tariff_tanzania[1:200,], insilico_tanzania[1:200,])
```

```{r ensemble_calibratedva}
ensemble_calibratedva <- calibratedva(A_U = ensemble_unlabeled,
                                      A_L = ensemble_labeled,
                                      G_L = gs_cod_tanzania[1:200,],
                                      causes = causes)
```

```{r ensemble_csmf}
ensemble_csmf_samples <- ggs(ensemble_calibratedva$samples, par_labels=P, family="p")
### CSMF with credible intervals
ensemble_csmf <-
  ensemble_csmf_samples %>%
  group_by(Parameter) %>%
  summarise(csmf = mean(value),
            ci_L = quantile(value, .025),
            ci_U = quantile(value, .975))
ensemble_csmf 
```

We can also view the traceplot

```{r ensemble_csmf_traceplot, fig.width = 8, fig.height = 12}
ggs_traceplot(ensemble_csmf_samples)
```

And a posterior density plot

```{r ensemble_csmf_density_plot, fig.width = 8, fig.height = 12}
ggs_density(ensemble_csmf_samples)
```

### Additional features

Please view the "Get Started" tab to see additional features in CalibratedVA.