---
title: "CalibratedVA"
author: "Jacob Fiksel and Abhirup Datta"
date: "`r Sys.Date()`"
output: pdf_document
bibliography: bibliography.bib
vignette: >
  %\VignetteIndexEntry{CalibratedVA}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  cache = TRUE
)
```

## Introduction

Many countries cannot conduct full autopsies for the majority of deaths due to either cultural or economic limitations [@abouzahr2015civil; @allotey2015let]. Rather than perform an invasive autopsy, an alternative method to determine the cause (or "etiology") of death (COD) is to conduct a *verbal autopsy* (VA), where relatives of the deceased individual are asked a set list of questions to obtain information about symptoms experienced by the deceased observed prior to death [@soleman2006verbal]. Medical experts can then
use these VA responses to determine the COD for an individual.

If we are interested in obtaining causes of death for a large amount of individuals in a
region or country, in order to determine optimal resource allocation, we need an efficient
way to scale the process of determining the COD from VA data. Several algorithms have
recently been proposed to automatically predict COD from VA records. These *computer coded* VA (CCVA) algorithms predict COD using the VA records as input. Examples include Tariff [@tariff; @tariff2], interVA [@interva], and InsilicoVA [@insilico]. These algorithms all require some gold standard (GS) information -- either a training dataset with both VA symptom data and
a gold standard (GS) COD determined for each individual or some symptom-given-cause probability matrix.

If we are interested in obtaining national estimates of cause specific mortality fraction (CSMF), we cantrain one of the above CCVA algorithms with GS COD data from our country of interest,
obtain individual predictions from CCVA data for the country of interest, and then aggregate the predictions to the CSMF level. However,
there may be limited amounts of GS COD data for a new country. In addition,
algorithms trained on GS COD from other countries may not be as accurate.

The CalibratedVA package is to be used for exactly the situation above--
we take predictions from one or multiple algorithm(s) trained on a large 
non-local dataset, and use the local GS COD dataset to improve the
CSMF estimates via local calibration.

## PHMRC Data

The Population Health Metrics Research Consortium (PHMRC) study contains GS COD data
for children, neonates and adults in 4 countries, making it a perfect dataset
to demonstrate the merits of `CalibratedVA`. 

We have used the openVA package to obtain cause-of-death predictions for adult
deaths from Tanzania. Both InSilicoVA and Tariff were trained using deaths
from the other 3 countries. We obtained individual cause probability predictions
from InSilicoVA and individual top cause-of-death predictions from Tariff. We
will also read in the GS-COD for each death from Tanzania.

```{r load_phmrc}
data("insilico_tanzania")
data("tariff_tanzania")
data("gs_cod_tanzania")
```

There are `r ncol(insilico_tanzania)` causes, which is too many causes for 
CalibratedVA to be useful. We will map these causes to broader causes, using the
following cause map.

```{r load_cause_map}
data("phmrc_adult_cause_map")
```

To map the causes, we can provide either the probability matrix:

```{r map_causes_insilicova}
insilico_tanzania <- map_causes(insilico_tanzania, phmrc_adult_cause_map)
```

Or the character vector with the predicted cause or GS-COD for each individual

```{r map_causes_insilicova}
tariff_tanzania <- map_causes(tariff_tanzania$cause, phmrc_adult_cause_map)
gs_cod_tanzania <- map_causes(gs_cod_tanzania, phmrc_adult_cause_map)
```

### CalibratedVA

#### Single algorithm predictions

We will first run CalibratedVA using the Tariff predictions. To run CalibratedVA, we will first select a set of individuals to be in our "hospital" set--that is, individuals for whom we know their GS-COD. These
individuals will be used for estimation of algorithm misclassification rates. We will
simply use the first 200 individuals from the data. We will use the default settings
for all hyperparameter values.


```{r tariff_calibratedva}
causes <- colnames(tariff_tanzania)
tariff_calibratedva <- calibratedva(A_U = tariff_tanzania[-(1:200),],
                                    A_L = tariff_tanzania[1:200,],
                                    G_L = gs_cod_tanzania[1:200,],
                                    causes = causes)
```

We can use the `ggmcmc` package to obtain CSMF estimates for the individual causes,
along with credible intervals.

```{r tariff_csmf}
library(ggmcmc)
P <- data.frame(
  Parameter=paste0("p[", 1:length(causes), "]"),
  Label=causes)
tariff_csmf_samples <- ggs(tariff_calibratedva$samples, par_labels=P, family="p")
### CSMF with credible intervals
tariff_csmf <-
  tariff_csmf_samples %>%
  group_by(Parameter) %>%
  summarise(csmf = mean(value),
            ci_L = quantile(value, .025),
            ci_U = quantile(value, .975))
tariff_csmf 
```

We can also view the traceplot

```{r tariff_csmf_traceplot, fig.width = 8, fig.height = 12}
ggs_traceplot(tariff_csmf_samples)
```

And a posterior density plot

```{r tariff_csmf_density_plot, fig.width = 8, fig.height = 12}
ggs_density(tariff_csmf_samples)
```

#### Using predictions from multiple algorithms (ensemble)

We can use both the predictions from both InSilicoVA and Tariff to 
better estimate the CSMF using the ensemble approach. We simply need to format
the predictions into a list of matrices

```{r format_for_ensemble}
ensemble_unlabeled <- list(tariff_tanzania[-(1:200),], insilico_tanzania[-(1:200),])
ensemble_labeled <- list(tariff_tanzania[1:200,], insilico_tanzania[1:200,])
```

```{r ensemble_calibratedva}
ensemble_calibratedva <- calibratedva(A_U = ensemble_unlabeled,
                                      A_L = ensemble_labeled,
                                      G_L = gs_cod_tanzania[1:200,],
                                      causes = causes)
```

```{r ensemble_csmf}
ensemble_csmf_samples <- ggs(ensemble_calibratedva$samples, par_labels=P, family="p")
### CSMF with credible intervals
ensemble_csmf <-
  ensemble_csmf_samples %>%
  group_by(Parameter) %>%
  summarise(csmf = mean(value),
            ci_L = quantile(value, .025),
            ci_U = quantile(value, .975))
ensemble_csmf 
```

We can also view the traceplot

```{r ensemble_csmf_traceplot, fig.width = 8, fig.height = 12}
ggs_traceplot(ensemble_csmf_samples)
```

And a posterior density plot

```{r ensemble_csmf_density_plot, fig.width = 8, fig.height = 12}
ggs_density(ensemble_csmf_samples)
```

### More details on shrinkage

#### Increasing levels of shrinkage

We can increase the alpha parameter to increase the amount of shrinkage. As an extreme 
example, we will set alpha = 500, and show that the estimated CSMF is very similar to the
predicted CSMF 

```{r tariff_calibratedva, cache = TRUE, message = FALSE}
tariff_highshrinkage <- calibratedva(A_U = tariff_tanzania[-(1:200),],
                                    A_L = tariff_tanzania[1:200,],
                                    G_L = gs_cod_tanzania[1:200,],
                                    causes = causes,
                                    alpha = 500)
```

```{r tariff_highshrinkage_csmf}
### CSMF with credible intervals
tariff_highshrinkage_csmf <-
  ggs(tariff_highshrinkage$samples, par_labels=P, family="p") %>%
  group_by(Parameter) %>%
  summarise(csmf = mean(value))
tariff_highshrinkage_csmf
colMeans(tariff_tanzania[-(1:200),])
```

#### Bayesian approach using CSMF shrinkage

Rather than shrinking the misclassification matrix ("M") to the identity matrix,
we can directly shrink the CSMF ("p") to the predicted CSMF on the unlabeled data.
This is simply done using `method = "pshrink"`. When using this method,
the level of shrinkage is controlled by `lambda`, where higher values
lead to more shrinkage. Setting `lambda = 1` (the default) will lead to little
shrinkage.

```{r tariff_pshrink, fig.width = 8, fig.height = 12}
tariff_pshrink <- calibratedva(A_U = tariff_tanzania[-(1:200),],
                               A_L = tariff_tanzania[1:200,],
                               G_L = gs_cod_tanzania[1:200,],
                               causes = causes,
                               method = "pshrink",
                               lambda = 1)
ggs_density(ggs(tariff_pshrink$samples, par_labels=P, family="p"))
```

We can see that the above approach leads to multimodality. We can use the 
`tune_calibratedva` function to choose a value of `lambda` which has the best WAIC,
and for which the posterior will be unimodal and will have a low Rhat. The below
example only uses 5 values of lambda for the gridsearch, but we recommend searching
over a larger number of values. We will also reduce the number of draws from 10,000 (default)
to 5,000. This is not necessarily recommended, but will reduce the run-time for 
this example.

```{r tariff_tune, fig.width = 8, fig.height = 12}
tariff_pshrink_tuned <- tune_calibratedva(A_U = tariff_tanzania[-(1:200),],
                                          A_L = tariff_tanzania[1:200,],
                                          G_L = gs_cod_tanzania[1:200,],
                                          causes = causes,
                                          method = "pshrink",
                                          lambda_vec  = c(.1, 1, 10, 50, 100),
                                          ndraws = 5000)
ggs_density(ggs(tariff_pshrink_tuned$final_model$samples, par_labels=P, family="p"))
```

We can also tune the models by collecting a list of models for different values
of lambda (or alpha for M-shrinkage), and supplying this list to `tune_calibratedva`

```{r tariff_tune}
lambda_vec <- c(.1, 1, 10, 50, 100)
tariff_samples_list <-lapply(lambda_vec, function(lambda) {
  calibrateva_out <- calibratedva(A_U, A_L, G_L, causes, method = "pshrink",
                                  lambda = lambda, ndraws = 5000)
  return(calibrateva_out)
})
tariff_pshrink_tuned2 <- tune_calibratedva(A_U = tariff_tanzania[-(1:200),],
                                           A_L = tariff_tanzania[1:200,],
                                           G_L = gs_cod_tanzania[1:200,],
                                           causes = causes,
                                           method = "pshrink",
                                           samples_list = tariff_samples_list,
                                           lambda_vec  = c(.1, 1, 10, 50, 100))
identical(tariff_pshrink_tuned$waic_df, tariff_pshrink_tuned2$waic_df)
```

## References
