---
title: "ReVAMP"
author: "Jacob Fiksel and Abhirup Datta"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

# Introduction

In many countries, it is infeasible to conduct full autopsies for each death, due to economic or infrastructural limitations. Consequently, national or sub-national statistics produced on cause-specific mortality numbers for many of these countries rely on data from verbal autopsy (VA) surveys. The VA data supplemented with physicians’ opinions or full autopsy reports for a smaller subset of the deaths are used to determine prediction rules that can ascertain a cause of death (COD) solely based on the VA report of an individual. Approaches to specify these rules include non-statistical algorithms including Tariff, InterVA and EAVA and more formal model based approaches like InsillicoVA that endows the prediction problem with a proper probabilistic framework allowing coherent statistical inference. To create the prediction algorithm, all of these approaches require a substantial training dataset, where both the VA reports and the true gold standard cause of deaths, either the full autopsies or the physicians' opinions are available.

The ReVAMP (Regularized Verbal Autopsy based Mortality Prediction) package is to be used
when a limited amount of gold standard cause of death data is available for a new country. 
Rather than train one of the existing algorithms on the extremely small gold standard data set, our algorithm uses a hierarchical Bayesian model in order to leverage the large amount of
PHMRC VA data available from other countries in informing CSMF estimates.

# Obtaining and splitting the PHMRC data

In this vignette, we will use the `openVA` package to obtain PHMRC data and implement 
the Tariff algorithm. We will of course also have to load the `ReVAMP` package.

```{r load_pkgs}
library(openVA)
library(ReVAMP)
```

We will now load in the PHMRC child data and use the `ConvertData.phmrc` function
to convert the data into the appropriate structure for Tariff.

```{r convert_data}
child.raw <- read.csv(getPHMRC_url("child"))
child.clean <- ConvertData.phmrc(child.raw, phmrc.type = "child")$output
```

For simplicity, we will restrict ourselves to the top 4 COD, and treat
all other COD as "other". Because "Other Defined Causes of Child Deaths" is
coded as cause 14, we will not include this cause in the Top 4.

```{r top_4_cod}
top.cod <- names(sort(table(child.clean$Cause), decreasing = TRUE))
top4.cod <- top.cod[top.cod != "14"][1:4]
child.clean$Cause[!(child.clean$Cause %in% top4.cod)] <- 99
```

We will now split the PHMRC data from India into a calibration and test set,
and use the rest of the PHMRC data as our training set. We will use a calibration
set size of 200.

```{r split_data}
set.seed(123)
countries <- ifelse(child.raw$site %in% c("AP", "UP"), "India", "Other")
india.data <- child.clean[countries == "India",]
train.data <- child.clean[countries == "Other",]
calibration.indices <- sample(nrow(india.data), 200, replace = F)
calibration.data <- india.data[calibration.indices,]
test.data <- india.data[-calibration.indices,]
```

# Implementing ReVAMP with Tariff

We will demonstrate how to use the ReVAMP method with the Tariff
training method for predicting COD from VA data. First we will train Tariff. 
Note that the `data` argument is the data for which we want to obtain COD
predictions for, which is both the calibration data and our test data.

```{r tariff_trainset}
set.seed(123)
tariff.fit <- codeVA(data = rbind(calibration.data, test.data),
                     data.type = "customize", model = "Tariff",
                     data.train = train.data, causes.train = "Cause")
```

We now need to extract three pieces of information for ReVAMP:

1. The individual COD predictions for the test set
2. The individual COD predictions for the calibration set
3. The gold-standard COD for the calibration set

Note that all of these must be character vectors

```{r extract_information_tariff}
pred.cod <- getTopCOD(tariff.fit)
pred.cod <- as.character(pred.cod[,2])
calib.cod <- pred.cod[1:nrow(calibration.data)]
test.cod <- pred.cod[-(1:nrow(calibration.data))]
calib.truth <- as.character(calibration.data$Cause)
```

We will now initiate hyper-parameter values for ReVAMP. These can be changed of course,
but we have found these values work well. We will set the number of draws
to 10,000 for computational speed in creating the Vignette, but we recommend
using more (50,000 - 150,000). Further more, we recommend running multiple
chains to check for convergence. This can be done by setting a new seed
before running the algorithm.

```{r hyperparams}
causes <- as.character(sort(unique(test.data$Cause)))
epsilon <- .001
alpha <- .001
beta <- .001
tau <- .1
tau.vec <- rep(tau, length(causes))
delta <- 1
gamma.init <- 1
ndraws <- 10E3
```

We will run the `revamp.sampler` using the output from the Tariff model. We will only
run one chain of the sampler in order to quickly compile the vignette.

```{r tariff_revamp}
set.seed(1234)
tariff.revamp <- revamp.sampler(test.cod = test.cod, calib.cod = calib.cod,
                                calib.truth = calib.truth, causes = causes,
                                epsilon = epsilon, alpha=alpha, beta=beta,
                                tau.vec=tau.vec, delta=delta,
                                gamma.init=gamma.init, ndraws = ndraws)
```


# Obtaining ReVAMP output

The `revamp.sampler` output is a list of length `ndraws`. Each element in this list is itself a list, with each element containing a posterior draw for each parameter. We can inspect what this looks like.

```{r inspect_posterior}
tariff.revamp[[5000]]
```

The most important parameters are the $p_{i}$, which is the true CSMF for COD $i$.
The $i$th element in $p_{i}$ refers to the $i$th element in the `causes` input
of `revamp.sampler`.

We can extract these parameter estimates and then plot the posterior densities versus
the true CSMF in the test set. We will use a burn-in of 2,000 and thin the draws by 5.

```{r post_dens_plot, fig.width = 6, fig.height = 6}
library(ggplot2)
library(dplyr)
burnin <- 2E3
thin <- 5

post.p <- lapply(seq(burnin, ndraws, by = thin), function(i) {
    draw.p <- tariff.revamp[[i]]$p
    return(data.frame(p = draw.p, cause = causes, draw = i))
})

post.p <- do.call(rbind, post.p)
true.p <- prop.table(table(test.data$Cause))
true.p <- data.frame(p = unname(as.vector(true.p)), cause = names(true.p))
ggplot() +
    geom_density(data = post.p, aes(p)) +
    geom_vline(data = true.p, aes(xintercept = p), colour = 'red') +
    facet_wrap(~cause)
```

We can also compare the CSMF Accuracy of the ReVAMP method to that of Tariff

```{r}
csmf.tariff <- sapply(causes, function(c) mean(test.cod == c))
csmf.revamp <- sapply(causes, function(c) mean(post.p$p[post.p$cause == c]))
csmf.test <- sapply(causes, function(c) mean(test.data$Cause == c))
tariff.csmf.acc <- getCSMF_accuracy(csmf.tariff, csmf.test)
revamp.csmf.acc <- getCSMF_accuracy(csmf.revamp, csmf.test)
```

# Ensemble ReVAMP

In practice, we may not want to rely on the COD estimated from just Tariff. Instead,
we may want to combine Tariff estimates with those of another algorithm such as
InsilicoVA. Furthermore, we can use the ReVAMP methodology to use both of these modeling
strategies.

We will first use InsilicoVA to get individual COD estimates for both our
calibration and test sets, as we did for Tariff.

```{r insilico_train}
insilico.train <- codeVA(data = rbind(calibration.data, test.data),
                         data.type = "customize", model = "InSilicoVA",
                         data.train = train.data, causes.train = "Cause",
                         jump.scale = 0.05, Nsim=10000, auto.length = FALSE)
```



# Implementing custom VA prediction methods

Instead of using the `revamp_with_va` function with the built-in VA prediction methods,
you may want to implement your own method. We will show how to do this, using the
Tariff method as an example.




### Formatting data for ReVAMP

We need two main inputs for ReVAMP. The first is a vector $v$ where $v_{i}$ is the
number of observations in the test (or test + calibration) set that the trained
(Tariff) model predicts to have died of cause $i$. The second is a matrix $T$ where 
$T_{i,j}$ is the number of records in the calibration set where the true COD
is $i$ and the predicted COD from the trained (Tariff) model is $j$. 

Note that $v$ should be of length $C$ and $T$ should be of dimension $C \times C$,
where $C$ is the total number of causes of death. Also, the $ith$ entry in $v$ should
correspond to the same COD as the $ith$ row (and column) in $T$. 

```{r format_for_revamp}
allcauses <- sort(unique(c(calibration.data$Cause, train.data$Cause)))
ncauses <- length(allcauses)
calibration.predictions <- all.predictions[1:nrow(calibration.data)]

v <- sapply(allcauses, function(c) sum(all.predictions == c))
names(v) <- allcauses
    
T.mat <- matrix(NA, nrow = ncauses, ncol = ncauses)
calib.truecod <- calibration.data$Cause
for(i in 1:ncauses){
    for(j in 1:ncauses){
        T.mat[i,j] <- sum(calib.truecod == allcauses[i] & calibration.predictions == allcauses[j])
    }
}
colnames(T.mat) <- rownames(T.mat) <- allcauses
```


And now we will run the ReVAMP algorithm, using the same hyperparameter values as before

```{r revamp.sampler}
set.seed(123)
posterior.custom <- revamp.sampler(v = v, T.mat = T.mat, epsilon = epsilon,
                               alpha=alpha, beta=beta, tau.vec=tau.vec, delta=delta,
                               gamma.init=gamma.init, ndraws = ndraws)
```

And we will confirm that this we obtained the same draws as with the `revamp_with_va` function

```{r check_equal_output}
identical(posterior.custom, posterior.chain1)
```

