---
title: "ReVAMP"
author: "Jacob Fiksel and Abhirup Datta"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = TRUE
)
```

## Introduction

In many countries, it is infeasible to conduct full autopsies for the majority of deaths due to economic and infrastructural limitations, and/or religious or cultural prohibitions against autopsies [@abouzahr2015civil; @allotey2015let]. An alternative method to infer the cause (or "etiology") of death (COD) is to conduct *verbal autopsy* (VA) --  a systematic interview of the relatives of the deceased individual to obtain information about symptoms observed prior to death [@soleman2006verbal]. The reported symptoms in the VA can then be assessed by a group of physicians who assign the likely cause(s) of death.

While *physician coded* VA is a useful alternative to full autopsies, scaling the process to the regional or national level is challenging. Consequently, over the last decade several data-driven solutions have been proposed that replace the physician coding with artificial intelligence software. These *computer coded* VA (CCVA) algorithms predict COD using the VA records as input. Examples include Tariff [@tariff; @tariff2], interVA [@interva], and InsilicoVA [@insilico]. To create the prediction algorithm, all of these approaches require a substantial training dataset, where both the VA reports and the true gold standard cause of deaths, either the full autopsies or the physicians' opinions are available.

The ReVAMP (Regularized Verbal Autopsy based Mortality Prediction) package is to be used
when a limited amount of gold standard (GS) cause of death data is available for a new country. 
ReVAMP implements a general calibration framework for obtaining population CSMF estimates in a new country using countrywide VA records and a limited amount of local GS data with the true COD. Our calibration approach is to first use one or more CCVA algorithms trained on non-local GS information, then use limited local GS data to update the initial CSMF to arrive at an improved final estimate.

## PHMRC Data

The Population Health Metrics Research Consortium (PHMRC) study, conducted in 4 countries across six sites, is a benchmark database of paired VA records and GS-COD of children, neonates and adults. PHMRC data is frequently used to assess performance of CCVA algorithms. In this vignette, we will use the `openVA` package to obtain PHMRC data and implement the Tariff and InSilicoVA algorithms. We will of course also have to load the `ReVAMP` package.


```{r load_pkgs}
library(openVA)
library(ReVAMP)
```

We will now load in the PHMRC child data and use the `ConvertData.phmrc` function
to convert the data into the appropriate structure for Tariff and InSilicoVA.

```{r convert_data}
child.raw <- read.csv(getPHMRC_url("child"))
child.clean <- ConvertData.phmrc(child.raw, phmrc.type = "child")$output
```

To demonstrate how to implement ReVAMP, we will first treat Tanzania as our country of interest
for obtaining accurate CSMF estimates. We will split the PHMRC data from Tanzania into a hospital (calibration) and population (test) set, and use the rest of the PHMRC data as our gold standard (training) set. We will use a calibration set size of 100.

```{r split_data}
countries <- ifelse(child.raw$site %in% c("Dar", "Pemba"), "Tanzania", "Other")
tanzania.data <- child.clean[countries == "Tanzania",]
train.data <- child.clean[countries == "Other",]
set.seed(851745)
calibration.indices <- sample(nrow(tanzania.data), 100, replace = F)
calibration.data <- tanzania.data[calibration.indices,]
test.data <- tanzania.data[-calibration.indices,]
```

## Implementing Tariff and InSilicoVA

In this section, we will use our training data to build prediction models using
the Tariff and InSilicoVA algorithms. First we will implement Tariff. 
Note that the `data` argument is the data for which we want to obtain COD
predictions for, which is both the calibration data and our test data (we 
will use the calibration set for ReVAMP later in the vignette). When evaluating
the performance of these algorithms, we will only do so on the predictions
for the test data.

```{r tariff_train}
set.seed(123)
tariff.train <- codeVA(data = rbind(calibration.data, test.data),
                     data.type = "customize", model = "Tariff",
                     data.train = train.data, causes.train = "Cause")
```

We will now implement InSilicoVA.

```{r insilico_train}
set.seed(123)
insilico.train <- codeVA(data = rbind(calibration.data, test.data),
                         data.type = "customize", model = "InSilicoVA",
                         data.train = train.data, causes.train = "Cause",
                         jump.scale = 0.05, Nsim=5000, auto.length = FALSE)
```


## Implementing ReVAMP with Individual Algorithms

For simplicity, we will restrict ourselves to predicting the distribution for the
top 3 COD (from Tanzania), and treat all other COD as "other". Because "Other Defined Causes of Child Deaths" is
coded as cause 14, we will not include this cause in the Top 4. 

We will extract predictions for the test and calibration set from both algorithms
implemented above, and then change any prediction that is not in the top 3 COD 
to other.

```{r top_3_cod}
top.cod <- names(sort(table(tanzania.data$Cause), decreasing = TRUE))
top3.cod <- top.cod[top.cod != "14"][1:3]
change.cause <- function(cause) {
    cause <- as.character(cause)
    cause[!(cause %in% top3.cod)] <- "99"
    return(cause)
}
tariff.train.cod <- change.cause(getTopCOD(tariff.train)[,2])
insilico.train.cod <- change.cause(getTopCOD(insilico.train)[,2])
test.changedcod <- change.cause(test.data$Cause)
calibration.changedcod <- change.cause(calibration.data$Cause)
```

Finally, we will separate out the InSilicoVA and Tariff predictions for
the test and calibration sets.

```{r separate_cod}
tariff.train.cod.test <- tariff.train.cod[-(1:100)]
tariff.train.cod.calib <- tariff.train.cod[1:100]
insilico.train.cod.test <- insilico.train.cod[-(1:100)]
insilico.train.cod.calib <- insilico.train.cod[1:100]
```

We will now initiate hyper-parameter values for ReVAMP. These can be changed of course,
but we have found these values work well. We will run 3  chains, obtaining 50,000 draws
for each chains. Note that the order of the causes in the `causes` vector, 
as this is the order in which ReVAMP will present the distribution of
COD estimates.

```{r hyperparams}
causes <- as.character(sort(unique(test.changedcod)))
epsilon <- .001
alpha <- .001
beta <- .001
tau <- .1
tau.vec <- rep(tau, length(causes))
delta <- 1
gamma.init <- 1
ndraws <- 50E3
nchains <- 3
```

We will first run ReVAMP using the Tariff predictions.

```{r tariff_revamp}
set.seed(123)
revamp.seeds <- sample(1e6, nchains, replace = F)
tariff.revamp <- lapply(1:nchains, function(i) {
    set.seed(revamp.seeds[i])
    revamp.sampler(test.cod = tariff.train.cod.test, calib.cod = tariff.train.cod.calib,
                   calib.truth = calibration.changedcod, causes = causes,
                   epsilon = epsilon, alpha=alpha, beta=beta,
                   tau.vec=tau.vec, delta=delta,
                   gamma.init=gamma.init, ndraws = ndraws)
})
```

And now using the InSilicoVA predictions.

```{r insilico_revamp}
set.seed(123)
revamp.seeds <- sample(1e6, nchains, replace = F)
insilico.revamp <- lapply(1:nchains, function(i) {
    set.seed(revamp.seeds[i])
    revamp.sampler(test.cod = insilico.train.cod.test, calib.cod = insilico.train.cod.calib,
                   calib.truth = calibration.changedcod, causes = causes,
                   epsilon = epsilon, alpha=alpha, beta=beta,
                   tau.vec=tau.vec, delta=delta,
                   gamma.init=gamma.init, ndraws = ndraws)
})
```

## Implementing Ensemble ReVAMP

The previous section clearly reveals that REVAMP can use the output from any CCVA algorithm. Several CCVA algorithms have been implemented in publicly available software; in any particular analysis, the optimal CCVA is unknown. For example, in the above implentation, if our population (test) set had not already been labeled with the GS COD, we would not know whether it would be best to implement ReVAMP with Tariff or InSilicoVA. In the code below, we show how to implement the Independent ReVAMP Ensemble Model, which uses the predictions from multiple VA-based algorithms. We use output from the Tariff and InSilicoVA models built in the previous section.

```{r revamp_ensemble}
set.seed(123)
revamp.seeds <- sample(1e6, nchains, replace = F)
test.cod.mat <- matrix(c(tariff.train.cod.test,  insilico.train.cod.test), ncol = 2)
calib.cod.mat <- matrix(c(tariff.train.cod.calib,  insilico.train.cod.calib), ncol = 2)
ensemble.revamp <- lapply(1:nchains, function(i) {
    set.seed(revamp.seeds[i])
    revamp.ensemble.lite.sampler(test.cod.mat = test.cod.mat, calib.cod.mat = calib.cod.mat,
                                 calib.truth = calibration.changedcod, causes = causes,
                                 epsilon = epsilon, alpha=alpha, beta=beta,
                                 tau.vec=tau.vec, delta=delta,
                                 gamma.init=gamma.init, ndraws = ndraws)
})
```

## Obtaining ReVAMP output

The `revamp.sampler` (or `revamp.ensemble.lite.sampler`) output is a list of length `ndraws`. Each element in this list is itself a list, with each element containing a posterior draw for each parameter. We can inspect what this looks like, from our first chain of the ReVAMP with Tariff sampler.

```{r inspect_posterior}
tariff.revamp[[1]][[5000]]
```

The most important parameters are the $p_{i}$, which is the true CSMF for COD $i$.
The $i$th element in $p_{i}$ refers to the $i$th element in the `causes` input
of `revamp.sampler`.

We can extract these parameter estimates using the `revampCSMF` function and then plot the posterior densities versus the true CSMF in the test set. We will use a burn-in of 10,000 and thin the draws by 10.

```{r extract_csmf}
burnin <- 10E3
thin <- 10

tariff.revamp.csmf.list <- lapply(1:nchains, function(i) {
    revampCSMF(tariff.revamp[[i]], burnin = burnin, thin = thin)
})
tariff.revamp.csmf.df <- do.call(rbind, tariff.revamp.csmf.list)

insilico.revamp.csmf.list <- lapply(1:nchains, function(i) {
    revampCSMF(insilico.revamp[[i]], burnin = burnin, thin = thin)
})
insilico.revamp.csmf.df <- do.call(rbind, insilico.revamp.csmf.list)

ensemble.revamp.csmf.list <- lapply(1:nchains, function(i) {
    revampCSMF(ensemble.revamp[[i]], burnin = burnin, thin = thin)
})
ensemble.revamp.csmf.df <- do.call(rbind, ensemble.revamp.csmf.list)

```

We can plot the posterior distribution of CSMF estimates from our ReVAMP with
Tariff sample. We will show the true CSMF in red, and the estimated CSMF
from just Tariff in blue.

```{r post_tariff_dens_plot, fig.width = 6, fig.height = 6}
library(ggplot2)
library(dplyr)

true.p <- prop.table(table(change.cause(tanzania.data$Cause)))
true.p <- data.frame(p = unname(as.vector(true.p)), cause = names(true.p))
tariff.p <- prop.table(table(tariff.train.cod.test))
tariff.p <- data.frame(p = unname(as.vector(tariff.p)), cause = names(tariff.p))
ggplot() +
    geom_density(data = tariff.revamp.csmf.df, aes(p)) +
    geom_vline(data = true.p, aes(xintercept = p), colour = 'red') +
    geom_vline(data = tariff.p, aes(xintercept = p), colour = 'blue') +
    facet_wrap(~cause, scales = "free_y") +
    xlim(0, 1)
```

## Comparing algorithm accuracy

A common metric for comparing CSMF predictions is CSMF Accuracy, which is defined as 

$$
CSMF_{acc} = 1 - \frac{\sum_{i=1}^{C} (CSMF_{i} - CSMF_{i}^{(true)})}{2(1 - \text{min}\{CSMF^{(true)}\})}
$$

where $CSMF_{i}$ and $CSMF_{i}^{(true)}$ are the estimated and true percentages, respectively, of the population for which the GS COD is cause $i$, $i=1, \ldots, C$.

To obtain estimates of the CSMFs, we will use the marginal posterior means from the 
posterior draws. We will first write functions to obtain the point predictions
from ReVAMP, Tariff, and InSilicoVA.

```{r csmf_functions}
revampCSMFMeanEstimate <- function(revamp.csmf.df, causes) {
    csmf.init <- sapply(causes, function(c) {
        revamp.cause <- revamp.csmf.df[revamp.csmf.df$cause == c,]
        return(mean(revamp.cause$p))
    })
    return(csmf.init)
}

openVACSMF <- function(topcod, causes) {
    csmf <- sapply(causes, function(c) mean(topcod == c))
    return(csmf)
}
```

We will now obtain these CSMF estimates, using the causes defined in the code 
block where we set the ReVAMP hyperparameter values.

```{r obtain_csmf}
tariff.revamp.csmf <- revampCSMFMeanEstimate(tariff.revamp.csmf.df, causes)
insilico.revamp.csmf <- revampCSMFMeanEstimate(insilico.revamp.csmf.df, causes)
ensemble.revamp.csmf <- revampCSMFMeanEstimate(ensemble.revamp.csmf.df, causes)
tariff.train.csmf <- openVACSMF(tariff.train.cod.test, causes)
insilico.train.csmf <- openVACSMF(insilico.train.cod.test, causes)
```

Finally, we will obtain and display the CSMF accuracy for these 5 methods.

```{r csmf_accuracy}
methods <- c("tariff_revamp",
             "insilico_revamp",
             "ensemble_revamp",
             "tariff_train",
             "insilico_train")
ptrue <- sapply(causes, function(c) mean(change.cause(tanzania.data$Cause) == c))
csmf.acc.df <- data.frame(csmf.acc = c(getCSMF_accuracy(tariff.revamp.csmf, ptrue),
                                       getCSMF_accuracy(insilico.revamp.csmf, ptrue),
                                       getCSMF_accuracy(ensemble.revamp.csmf, ptrue),
                                       getCSMF_accuracy(tariff.train.csmf, ptrue),
                                       getCSMF_accuracy(insilico.train.csmf, ptrue)),
                          method = methods)
csmf.acc.df
```

We see two interesting observations. First, all three ReVAMP methods
substantially outperform Tariff and InSilicoVA in terms of CSMF accuracy. Second,
The CSMF accuracy of the ensemble method of ReVAMP falls right in the middle of 
the CSMF accuracies of Tariff and ReVAMP. This indicates that it is a useful method
to use, as it is always unknown whether Tariff or InSilicoVA will be the best performing algorithm.

## References
